An Empirical Analysis of the Internet Engineering Task Force with Computational Methods

### 2024-09-21 paper.tex
Updates to Trends in RFC Publication

%--------------------------------------------------------------------------------------------------
\subsection{Document Production and Complexity}

@@

\pb{Growth of RFCs.}
In total, 8,711 RFCs have been published through to the end of 2020.
Figure~\ref{fig:rfcs_by_area} shows how publication trends, in terms of
IETF areas and non-IETF streams, have changed over time. We identify three
broad publication phases in the RFC series. First, in 1969 through 1974,
RFCs are published at a rapid rate during the initial development of the
ARPANET. Then, from 1975 through 1985, development slows. This reflects a
relatively small community that is gaining real-world experience with the
network and developing a small core of applications and protocols. Finally,
with the creation of the IETF and the introduction of the National Science
Foundation Network (NSFNET) in 1986, both the community, and the number of
RFCs published, starts to expand rapidly. This is further driven by the
opening of the network to commercial and public use in the mid-1990s, and
continues to this day.

@@

\pb{Role of Working Groups (WGs).}
From the creation of the IETF in 1986, the growing community with its interests
in an increasing set of applications and protocols, has been split into working
groups. These working groups are chartered with a focus on a well-defined
programme of work, and exist within areas that have a broader focus.
As shown in
Figure~\ref{fig:rfcs_by_area}, the output of different areas has remained
relatively stable over time. The most notable trends begin with the creation of
the Real-time Applications and Infrastructure (rai) area from within the
Transport (tsv) area, and its later merger with the Applications (app) area to
become the Applications and Real-Time (art) area around 2014. Additionally, we also observe the significant growth in output of the Routing (rtg) area,
owing to the ongoing development of standards for MPLS, service function
chaining, and fat tree routing in data centres.

@@

%--------------------------------------------------------------------------------------------------
\subsection{Document Errata}

>>>>

%--------------------------------------------------------------------------------------------------
\subsection{RFC Production and Complexity}

RFC production trends show clear evidence that the IETF is a mature
standards development organisation. The peak of activity was in the 1990s
and early 2000s, with the initial broad deployment of the Internet, and the
organisation has since entered more of a development and maintenance mode
of operation. There are clear signs that the complexity of the installed
base of protocols is slowing development of new technologies.

@@

\pb{RFC Publication Rate:}
In total, 8,711 RFCs have been published through to the end of 2020.
Figure~\ref{fig:rfcs_by_area} shows how publication trends, in terms of
IETF areas and non-IETF streams, have changed over time. We identify three
broad publication phases in the RFC series. First, in 1969 through 1974,
RFCs are published at a rapid rate during the initial development of the
ARPANET. Then, from 1975 through 1985, development slows. This reflects a
relatively small community that is gaining real-world experience with the
network and developing a small core of applications and protocols. Finally,
with the creation of the IETF and the introduction of the National Science
Foundation Network (NSFNET) in 1986, both the community, and the number of
RFCs published, starts to expand rapidly. This is further driven by the
opening of the network to commercial and public use in the mid-1990s, and
continues to this day.

@@

\pb{Role of Working Groups:}
From the creation of the IETF in 1986, the growing community with its
interests in an increasing set of applications and protocols, has been
split into working groups. These working groups are chartered with a focus
on a well-defined programme of work, and exist within areas that have a
broader focus.  As shown in Figure~\ref{fig:rfcs_by_area}, the output of
different areas has remained relatively stable over time. The most notable
trends begin with the creation of the Real-time Applications and
Infrastructure (rai) area from within the Transport (tsv) area, and its
later merger with the Applications (app) area to become the Applications
and Real-Time (art) area around 2014. Additionally, we also observe the
significant growth in output of the Routing (rtg) area, owing to the
ongoing development of standards for MPLS, service function chaining, and
fat tree routing in data centres.

@@

%--------------------------------------------------------------------------------------------------
\subsection{RFC Errata}
